### 2020-03-13 Paper_EV/Paper_EV.tex
Backup 13_03_2020

The dynamic equations which describe the general motion of an aircraft are usually coupled first order implicit nonlinear differential equations. However, considering a quadrotor control problem a lot of plausible assumptions (e.g. neglecting the Coriolis and centrifugal acceleration due to the earth rotation) can be made to obtain more simplified versions of those equations. This section presents simplified general \acrshort{EOM} and describes the assumptions which were considered to derive them.

@@

\subsection{Inner INDI loop for Quadrotor} \label{subsec: indi_inner}

In this subsection, using theory from \ref{subsec:indi_general} the inner loop \acrshort{INDI} controller is derived to control angular acceleration of the Crazyflie quadrotor. The derivation of the inner loop, as well as the outer loop controllers is based on the \acrshort{INDI} controller architecture introduced by Smeur \cite{Smeur1}, \cite{Smeur2}. Equation \ref{eq:ang_momentum} from section \ref{sec:eqs_motion} serves as a basis for this derivation. The desired variable to be controlled by the inner loop \acrshort{INDI} is the angular acceleration of the quadrotor in the body-fixed coordinate frame. As in the case of a real quadrotor control problem the value of the thrust can be seen as an output of the dynamic system, it makes sense to incorporate thrust as a control variable into the control law as well. Thus, the angular momentum Equation \ref{eq:ang_momentum} is augmented with the total thrust $T$ of all four rotors \cite{Smeur2}. As only the body-fixed frame is used in the following control law derivation, the subscripts $B$ are not used in this subsection. Also all external forces and moments apply to the center of gravity of the quadrotor, thus the superscripts $G$ are also omitted. 
Solving Equation \ref{eq:ang_momentum} for the angular acceleration results in
\begin{equation}
	\begin{bmatrix}
		\dot{\bm{\omega}}\\
		T
	\end{bmatrix} = 
	\underbrace{
	\begin{bmatrix}
		-\bm{I}^{-1} \big(\bm{\omega} \times \bm{I} \bm{\omega} \big)\\
		0
	\end{bmatrix}}_\text{$\bm{F}(\bm{\omega})$} +
	\underbrace{	
	\begin{bmatrix}
		\bm{I}^{-1} \big( \bm{M}_G + \bm{M}_{A} +  \bm{M}_{P} \big)\\
		T
	\end{bmatrix}}_\text{$\bm{G}(\bm{\omega}, \bm{\Omega}, \bm{\dot{\Omega}})$}
	\label{eq:main_eom_indi_inner}
\end{equation}
where $\bm{\omega}$ and $\dot{\bm{\omega}}$ are angular velocity and acceleration of the body-fixed frame ($B$) with respect to the \acrshort{NED} ($O$) coordinate frame, $\bm{M}_G$ the gravitational moment, $\bm{M}_A$ the aerodynamic moment and $\bm{M}_P$ the propulsion moment. The vector $\bm{\Omega}$ contains angular velocities of all four rotors and serves as an input variable of the system. It is assumed that the gravitational force is applied to the center of gravity of the quadrotor and does not cause any moment around it. Due to the absence of the aerodynamic moment this term is also omitted and can be seen as a disturbance \cite{Smeur1}. The remaining propulsion moment is written as $\bm{M}_P=\bm{M}_C-\bm{M}_{gyro}$ where $\bm{M}_C$ is the control moment generated by the rotors and $\bm{M}_{gyro}$ the moment containing the gyroscopic effect of the rotors. This two moments can be explicitly written as
\begin{equation}
	\bm{M}_C = 	
	\begin{bmatrix}
		-bk_F & bk_F & bk_F & -bk_F \\
		lk_F & lk_F & -lk_F & -lk_F \\
		k_M & -k_M & k_M & -k_M 
	\end{bmatrix} \bm{\Omega}
	\label{eq:m_c}
\end{equation}
\begin{equation}
	\begin{split}
	\bm{M}_{gyro} = 
	\begin{bmatrix}
		0 & 0 & 0 & 0 \\
		0 & 0 & 0 & 0 \\
		I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz}
	\end{bmatrix} \bm{\dot{\Omega}}	+
	\begin{bmatrix}
		\bm{\omega}_y & 0 & 0 \\
		0 & \bm{\omega}_x & 0 \\
		0 & 0 & 0 
	\end{bmatrix}	
	\begin{bmatrix}
		I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz} \\
		-I_{rzz} & I_{rzz} & -I_{rzz} & I_{rzz} \\
		0 & 0 & 0 & 0 
	\end{bmatrix} \bm{\omega}
	\label{eq:m_gyro}
	\end{split}
\end{equation}
where $I_{rzz}$ is the element of the inertia matrix $I_{r}$ of the rotor, $l$ and $b$ lever arms as denoted in the Figure \ref{fig:frames_leverarms}, $k_F$ and $k_M$ force and moment constants of the rotors.

\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=1\textwidth]{figs/frames.pdf}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Crazyflie 2.1 nano-quadrotor with indicated lever arms of the motors]{Image of the Crazyflie 2.1 quadrotor adopted from \cite{bitcraze} with indicated lever arms of the motors.}
	\label{fig:frames_leverarms}
\end{figure}

As already explained in the previous subsection to derive the incremental control law the Taylor series expansion is perfromed on the Equation \ref{eq:main_eom_indi_inner}
\begin{equation}
	\begin{split}
		\begin{bmatrix}
			\bm{\dot{\omega}}\\
			T
		\end{bmatrix} &= \bm{F}(\bm{\omega}_0) + \bm{G}(\bm{\omega}_0, \bm{\Omega}_0, \bm{\dot{\Omega}}_0) \\
		&+ \frac{\partial}{\partial \bm{\omega}} \big[\bm{F}(\bm{\omega}) + \bm{G}(\bm{\omega}_0, \bm{\Omega}_0, \bm{\dot{\Omega}}_0) \big] \bigg| _{\bm{\omega}=\bm{\omega}_0} (\bm{\omega}-\bm{\omega}_0) \\
		&+ \frac{\partial}{\partial \bm{\Omega}} \big[\bm{G}(\bm{\omega}_0, \bm{\Omega}, \bm{\dot{\Omega}}_0) \big] \bigg| _{\bm{\Omega}=\bm{\Omega}_0} (\bm{\Omega}-\bm{\Omega}_0) \\
		&+ \frac{\partial}{\partial \bm{\dot{\Omega}}} \big[\bm{G}(\bm{\omega}_0, \bm{\Omega}_0, \bm{\dot{\Omega}}) \big] \bigg| _{\bm{\dot{\Omega}}=\bm{\dot{\Omega}}_0} (\bm{\dot{\Omega}}-\bm{\dot{\Omega}}_0)
		\label{eq:main_eom_indi_inner_taylor}
	\end{split}
\end{equation}
By applying differentiation and rewriting some of the terms, following equation is obtained
\begin{equation}
	\begin{bmatrix}
		\bm{\dot{\omega}}\\
		T
	\end{bmatrix} = 
	\begin{bmatrix}
		\bm{\dot{\omega}}_0\\
		T_0
	\end{bmatrix} + \bm{G}_1(\bm{\Omega} - \bm{\Omega}_0) + T_S \bm{G}_2(\bm{\dot{\Omega}} - \bm{\dot{\Omega}}_0)
	\label{eq:main_eom_indi_inner_lin}
\end{equation}
The first term of the Equation \ref{eq:main_eom_indi_inner_lin} is the angular acceleration based on the current angular rates $\bm{\omega}_0$ and inputs $\bm{\Omega}_0$ and can be denoted as $\bm{\dot{\omega}}_0$. $T_0$ is the current thrust value. The last term on the right side of the Equation \ref{eq:main_eom_indi_inner_lin} is scaled with the sample time $T_s$ which is introduced only to simplify further mathematical transformations. The expressions of the control moment $\bm{M}_C$ and the gyroscopic moment of the rotors $\bm{M}_{gyro}$ have been summarized to control effectiveness matrices $\bm{G}_1$ and $\bm{G}_2$
\begin{equation}
	\bm{G}_1 = 
	\begin{bmatrix}
		\bm{I}^{-1}\begin{bmatrix}
			-bk_F & bk_F & bk_F & -bk_F \\
			lk_F & lk_F & -lk_F & -lk_F \\
			k_M & -k_M & k_M & -k_M 
		\end{bmatrix} - 	
		\begin{bmatrix}
			\bm{\omega}_y & 0 & 0 \\
			0 & \bm{\omega}_x & 0 \\
			0 & 0 & 0 
		\end{bmatrix}	
		\begin{bmatrix}
			I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz} \\
			-I_{rzz} & I_{rzz} & -I_{rzz} & I_{rzz} \\
			0 & 0 & 0 & 0 
		\end{bmatrix}\\
			k_F \cdot \bm{1}_{1\times4}
	\end{bmatrix}
	\label{eq:G1}
\end{equation}
\begin{equation}
	\bm{G}_2 = 
	\begin{bmatrix}
		-\bm{I}^{-1} T_S^{-1}\begin{bmatrix}
			0 & 0 & 0 & 0 \\
			0 & 0 & 0 & 0 \\
			I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz}
		\end{bmatrix} \\
		\bm{0}_{1\times4}
	\end{bmatrix}
	\label{eq:G2}
\end{equation}
with the terms $\bm{1}_{1\times4}$ and $\bm{0}_{1\times4}$ being $1\times4$ vectors of ones and zeros, respectively.

To prepare the linearized dynamic Equation \ref{eq:main_eom_indi_inner_lin} for discrete implementation on a computing system, the discrete approximation ($z$ domain) of the derivative is used: $\bm{\dot{\Omega}} = (\bm{\Omega} - \bm{\Omega}z^{-1})T_S^{-1}$. Furthermore, angular acceleration $\bm{\dot{\omega}}_0$ which is obtained from the differentiated gyroscope measurements is usually noisy. It has been shown, that using a second order filtering can help to reduce measurement nose. At the same time such a filter introduces time delay, which has to be considered in the derivation, as it is important to have a unique delay for all variables which are used in the Taylor expansion \cite{Smeur1}. Thus, the same second order filter is applied to all variables with a subscript $0$. By applying filtering (subscript $f$) and finite differences method to the Equation \ref{eq:main_eom_indi_inner_lin}, its discrete version results in
\begin{equation}
	\begin{bmatrix}
		\bm{\dot{\omega}}\\
		T
	\end{bmatrix} = 
	\begin{bmatrix}
		\bm{\dot{\omega}}_f\\
		T_f
	\end{bmatrix} + (\bm{G}_1+\bm{G}_2)(\bm{\Omega} - \bm{\Omega}_f) - \bm{G}_2z^{-1}(\bm{\Omega} - \bm{\Omega}_f)
	\label{eq:main_eom_indi_inner_lin_final}
\end{equation}
By solving Equation \ref{eq:main_eom_indi_inner_lin_final} for $\bm{\dot{\Omega}}$ and substituting $\bm{\dot{\omega}}$ with the equivalent input $\bm{\nu}_{ang}$, the control law of the inner loop is obtained
\begin{equation}
	\begin{split}
		\bm{\dot{\Omega}}_c = \bm{\dot{\Omega}} = \bm{\dot{\Omega}}_f + (\bm{G}_1+\bm{G}_2)^{-1} \Bigg(
		\begin{bmatrix}
			\bm{\nu}_{ang} - \bm{\dot{\omega}}_f\\
			\tilde{T}
		\end{bmatrix} + \bm{G}_2z^{-1}(\bm{\Omega} - \bm{\Omega}_f) \Bigg)
		\label{eq:indi_inner_control_law}	
	\end{split}
\end{equation}
where $\bm{\dot{\Omega}}$ is a vector of commanded rotational rates for every rotor and $\tilde{T} = T - T_f$ being the thrust increment which is provided by the outer loop \acrshort{INDI}. The block diagram of the inner loop \acrshort{INDI} controller is presented in Figure (TODO:Figure). In the figure the derived controller which controls the angular acceleration is augmented with two simple controllers which control the angular velocity $\bm{\omega}$ and the attitude $\bm{\eta}$ of the quadrotor. Each controller consists of a single gain. Thus, the reference values of the \acrshort{INDI} part of the inner loop are provided by these two controllers.

\subsection{Outer INDI loop for Quadrotor} \label{subsec: indi_outer}

@@

The linear acceleration $\bm{\dot{v}}_0$ which is obtained from accelerometer measurements is usually noisy. Applying the same filter as in the case of the inner loop derivation, the Equation \ref{eq:lin_momentum_indi_outer_lin} is given by 
\begin{equation}
	\begin{split}
		\bm{\dot{v}} = \bm{\dot{v}}_f +  m^{-1} \bm{G}(\Phi_0, \Theta_0, \Psi_0, T_0) (\bm{u} - \bm{u}_f)
		\label{eq:lin_momentum_indi_outer_lin_final}
	\end{split}
\end{equation}
Solving Equation \ref{eq:lin_momentum_indi_outer_lin_final} for $\bm{u}$ and substituting $\bm{\dot{v}}$ with the equivalent control variable $\bm{\nu}_{lin}$, the control law of the outer loop is obtained
\begin{equation}
	\begin{split}
		\bm{u_c} = \bm{u} = \bm{u}_f +  m \bm{G}^{-1}(\Phi_0, \Theta_0, \Psi_0, T_0) (\bm{\nu}_{lin} - \bm{\dot{v}}_f)
		\label{eq:indi_outer_control_law}
	\end{split}
\end{equation}
Figure (TODO:show Figure) shows the block diagram of the outer loop \acrshort{INDI}. Using the same approach as with the inner loop, the outer loop which controls linear acceleration is augmented with two closed loops consisting of two gains. These gains represent two controllers which control linear velocity and position.

TODOs: \\
- INDI inner loop (order)\\
- 2.3.2 Figure with Crazyflie lever arms and motor numbers \\
- 2.3.2 Figure with block diagram of the inner loop controller\\
- 2.3.3 Figure with block diagram of the outer loop controller

@@

The quadrotor used in this project was extended with a flow deck sensor. The flow deck itself is equipped with two sensors: the optical flow sensor which measures the visual motion and the \acrfull{ToF} sensor which measures the range to the next object along the body-fixed z-axis of the quadcopter. By employing these sensors, the flow deck can provide estimates of the current position and linear velocity.

\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=0.4\textwidth]{figs/crazyflie_drone.jpg}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Crazyflie 2.1 nano-quadrotor]{Crazyflie 2.1 nano-quadrotor \cite{bitcraze}}
	\label{fig:crazyflie_drone}
\end{figure}

@@

In the framework of this thesis the Simulink model was also applied to better understand the controller structure and its characteristics. Based on the simulation it was determined which of the control effectiveness terms have a greater influence on the system response and which ones can be neglected. Additionally the simulation model provided a good overview of potential computational problems and mistakes which could lead to drastically growing values and unstable behaviour of the system. To solve some of them, saturation blocks have been used  in different places.

@@

The core of the Crazyflie simulation is the model of the quadrotor dynamics. It is based on dynamic equations of motion presented in Chapter \ref{sec:eqs_motion}. Similarly to the section \ref{subsec: indi_outer}, it was assumed that the sum of the forces in the linear momentum equation is composed of three components, namely gravitational, propulsive and aerodynamic. The gravity and the propulsive force have not been changed, so the Equations \ref{eq:Fg} and \ref{eq:Fp} were also adopted for the simulation model. The thrust value $T$ was modelled as a linear function of rotational rates of propellers 
\begin{equation}
	\begin{split}
		T = - k_F (\Omega_1 + \Omega_2 + \Omega_3 + \Omega_4)
		\label{eq:thrust}
	\end{split}
\end{equation}
where $k_F$ is the force constant of the rotors. In subsection \ref{subsec:parameter_estimation} it is shown how to estimate the thrust coefficient directly from the flight data. However, for the simulation model $k_F$ was computed based on the assumption that the influence of the force constant of the rotors $k_F$ (considering the lever arm) is greater than the influence of the moment constant of the rotors $k_M$, and that the thrust to weight ratio expression $\frac{T}{W} \approx 2$.

@@

\subsection{Testing with contact Forces and Moments?} \label{subsec:testing_with_fandm}

@@

- To make the controller work a minimal knowledge of the system dynymics is need. Nevertherless G1, G2 still have to estimated accuratly because this 2 parameters do have inpact on the controller performance. \\
- Plain step responses of the outer loop \\
- Step responses with disturbance of the outer loop

>>>>

The dynamic equations which describe the general motion of an aircraft are usually coupled first order implicit nonlinear differential equations. However, considering a quadrotor control problem, a lot of plausible assumptions (e.g. neglecting the Coriolis and centrifugal acceleration due to the earth rotation) can be made to obtain more simplified versions of those equations. This section presents simplified general \acrshort{EOM} and describes the assumptions which were considered to derive them.

@@

\subsection{Inner loop INDI for Quadrotor} \label{subsec: indi_inner}

In this subsection, using theory from \ref{subsec:indi_general} the inner loop \acrshort{INDI} controller is derived to control angular acceleration of the Crazyflie quadrotor. The derivation of the inner loop, as well as the outer loop controllers is based on the \acrshort{INDI} controller architecture introduced by Smeur \cite{Smeur1}, \cite{Smeur2}. Equation \ref{eq:ang_momentum} from section \ref{sec:eqs_motion} serves as a basis for this derivation. The desired variable to be controlled by the inner loop \acrshort{INDI} is the angular acceleration of the quadrotor in the body-fixed coordinate frame. As in the case of a real quadrotor control problem the value of the thrust can be seen as an output of the dynamic system, it makes sense to incorporate thrust as a control variable into the control law as well. Thus, the angular momentum Equation \ref{eq:ang_momentum} is augmented with the total thrust $T$ of all four rotors \cite{Smeur2}. As only the body-fixed frame is used in the following control law derivation, the subscripts $B$ are not used in this subsection. Also all external forces and moments apply to the center of gravity of the quadrotor, thus the superscripts $G$ are also omitted. 
Solving Equation \ref{eq:ang_momentum} for the angular acceleration results in
		\dot{\bm{\omega}}\\
	\end{bmatrix} = 
		-\bm{I}^{-1} \big(\bm{\omega} \times \bm{I} \bm{\omega} \big)\\
		0
	\end{bmatrix}}_\text{$\bm{F}(\bm{\omega})$} +
		\bm{I}^{-1} \big( \bm{M}_G + \bm{M}_{A} +  \bm{M}_{P} \big)\\
	\end{bmatrix}}_\text{$\bm{G}(\bm{\omega}, \bm{\Omega}, \bm{\dot{\Omega}})$}
	\label{eq:main_eom_indi_inner}
where $\bm{\omega}$ and $\dot{\bm{\omega}}$ are angular velocity and acceleration of the body-fixed frame ($B$) with respect to the \acrshort{NED} ($O$) coordinate frame, $\bm{M}_G$ the gravitational moment, $\bm{M}_A$ the aerodynamic moment and $\bm{M}_P$ the propulsion moment. The vector $\bm{\Omega}$ contains angular velocities of all four rotors and serves as an input variable of the system. It is assumed that the gravitational force is applied to the center of gravity of the quadrotor and does not cause any moment around it. Due to the absence of the aerodynamic moment this term is also omitted and can be seen as a disturbance \cite{Smeur1}. The remaining propulsion moment is written as $\bm{M}_P=\bm{M}_C-\bm{M}_{gyro}$ where $\bm{M}_C$ is the control moment generated by the rotors and $\bm{M}_{gyro}$ the moment containing the gyroscopic effect of the rotors. This two moments can be explicitly written as
	\bm{M}_C = 	
		-bk_F & bk_F & bk_F & -bk_F \\
		lk_F & lk_F & -lk_F & -lk_F \\
		k_M & -k_M & k_M & -k_M 
	\end{bmatrix} \bm{\Omega}
	\label{eq:m_c}
	\bm{M}_{gyro} = 
		0 & 0 & 0 & 0 \\
		0 & 0 & 0 & 0 \\
		I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz}
	\end{bmatrix} \bm{\dot{\Omega}}	+
		\bm{\omega}_y & 0 & 0 \\
		0 & \bm{\omega}_x & 0 \\
		0 & 0 & 0 
		I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz} \\
		-I_{rzz} & I_{rzz} & -I_{rzz} & I_{rzz} \\
		0 & 0 & 0 & 0 
	\end{bmatrix} \bm{\omega}
	\label{eq:m_gyro}
where $I_{rzz}$ is the element of the inertia matrix $I_{r}$ of the rotor, $l$ and $b$ lever arms as denoted in the Figure \ref{fig:frames_leverarms}, $k_F$ and $k_M$ force and moment constants of the rotors.
\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=0.8\textwidth]{figs/frames.pdf}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Crazyflie 2.1 nano-quadrotor with indicated lever arms of the motors]{Image of the Crazyflie 2.1 quadrotor adopted from \cite{bitcraze} with indicated lever arms of the motors.}
	\label{fig:frames_leverarms}
\end{figure}
As already explained in the previous subsection to derive the incremental control law the Taylor series expansion is perfromed on the Equation \ref{eq:main_eom_indi_inner}
			\bm{\dot{\omega}}\\
		\end{bmatrix} &= \bm{F}(\bm{\omega}_0) + \bm{G}(\bm{\omega}_0, \bm{\Omega}_0, \bm{\dot{\Omega}}_0) \\
		&+ \frac{\partial}{\partial \bm{\omega}} \big[\bm{F}(\bm{\omega}) + \bm{G}(\bm{\omega}_0, \bm{\Omega}_0, \bm{\dot{\Omega}}_0) \big] \bigg| _{\bm{\omega}=\bm{\omega}_0} (\bm{\omega}-\bm{\omega}_0) \\
		&+ \frac{\partial}{\partial \bm{\Omega}} \big[\bm{G}(\bm{\omega}_0, \bm{\Omega}, \bm{\dot{\Omega}}_0) \big] \bigg| _{\bm{\Omega}=\bm{\Omega}_0} (\bm{\Omega}-\bm{\Omega}_0) \\
		&+ \frac{\partial}{\partial \bm{\dot{\Omega}}} \big[\bm{G}(\bm{\omega}_0, \bm{\Omega}_0, \bm{\dot{\Omega}}) \big] \bigg| _{\bm{\dot{\Omega}}=\bm{\dot{\Omega}}_0} (\bm{\dot{\Omega}}-\bm{\dot{\Omega}}_0)
		\label{eq:main_eom_indi_inner_taylor}
By applying differentiation and rewriting some of the terms, following equation is obtained
		\bm{\dot{\omega}}\\
	\end{bmatrix} = 
		\bm{\dot{\omega}}_0\\
		T_0
	\end{bmatrix} + \bm{G}_1(\bm{\Omega} - \bm{\Omega}_0) + T_s \bm{G}_2(\bm{\dot{\Omega}} - \bm{\dot{\Omega}}_0)
	\label{eq:main_eom_indi_inner_lin}
The first term of the Equation \ref{eq:main_eom_indi_inner_lin} is the angular acceleration based on the current angular rates $\bm{\omega}_0$ and inputs $\bm{\Omega}_0$ and can be denoted as $\bm{\dot{\omega}}_0$. $T_0$ is the current thrust value. The last term on the right side of the Equation \ref{eq:main_eom_indi_inner_lin} is scaled with the sample time $T_s$ which is introduced only to simplify further mathematical transformations. The expressions of the control moment $\bm{M}_C$ and the gyroscopic moment of the rotors $\bm{M}_{gyro}$ have been summarized to control effectiveness matrices $\bm{G}_1$ and $\bm{G}_2$
	\bm{G}_1 = 
		\bm{I}^{-1}\begin{bmatrix}
			-bk_F & bk_F & bk_F & -bk_F \\
			lk_F & lk_F & -lk_F & -lk_F \\
			k_M & -k_M & k_M & -k_M 
		\end{bmatrix} - 	
		\bm{I}^{-1}\begin{bmatrix}
			\bm{\omega}_y & 0 & 0 \\
			0 & \bm{\omega}_x & 0 \\
			0 & 0 & 0 
			I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz} \\
			-I_{rzz} & I_{rzz} & -I_{rzz} & I_{rzz} \\
			0 & 0 & 0 & 0 
		\end{bmatrix}\\
			k_F \cdot \bm{1}_{1\times4}
	\label{eq:G1}
	\bm{G}_2 = 
		-\bm{I}^{-1} T_s^{-1}\begin{bmatrix}
			0 & 0 & 0 & 0 \\
			0 & 0 & 0 & 0 \\
			I_{rzz} & -I_{rzz} & I_{rzz} & -I_{rzz}
		\end{bmatrix} \\
		\bm{0}_{1\times4}
	\label{eq:G2}
with the terms $\bm{1}_{1\times4}$ and $\bm{0}_{1\times4}$ being $1\times4$ vectors of ones and zeros, respectively.

To prepare the linearized dynamic Equation \ref{eq:main_eom_indi_inner_lin} for discrete implementation on a computing system, the discrete approximation ($z$ domain) of the derivative is used: $\bm{\dot{\Omega}} = (\bm{\Omega} - \bm{\Omega}z^{-1})T_s^{-1}$. Furthermore, angular acceleration $\bm{\dot{\omega}}_0$ which is obtained from the differentiated gyroscope measurements is usually noisy. It has been shown, that using a second order filtering can help to reduce measurement nose. At the same time such a filter introduces time delay, which has to be considered in the derivation, as it is important to have a unique delay for all variables which are used in the Taylor expansion \cite{Smeur1}. Thus, the same second order filter is applied to all variables with a subscript $0$. By applying filtering (subscript $f$) and finite differences method to the Equation \ref{eq:main_eom_indi_inner_lin}, its discrete version results in
\begin{equation}
	\begin{bmatrix}
		\bm{\dot{\omega}}\\
		T
	\end{bmatrix} = 
	\begin{bmatrix}
		\bm{\dot{\omega}}_f\\
		T_f
	\end{bmatrix} + (\bm{G}_1+\bm{G}_2)(\bm{\Omega} - \bm{\Omega}_f) - \bm{G}_2z^{-1}(\bm{\Omega} - \bm{\Omega}_f)
	\label{eq:main_eom_indi_inner_lin_final}
\end{equation}
By solving Equation \ref{eq:main_eom_indi_inner_lin_final} for $\bm{\dot{\Omega}}$ and substituting $\bm{\dot{\omega}}$ with the equivalent input $\bm{\nu}_{ang}$, the control law of the inner loop is obtained
\begin{equation}
	\begin{split}
		\bm{\dot{\Omega}}_c = \bm{\dot{\Omega}} = \bm{\dot{\Omega}}_f + (\bm{G}_1+\bm{G}_2)^{-1} \Bigg(
		\begin{bmatrix}
			\bm{\nu}_{ang} - \bm{\dot{\omega}}_f\\
			\tilde{T}
		\end{bmatrix} + \bm{G}_2z^{-1}(\bm{\Omega} - \bm{\Omega}_f) \Bigg)
		\label{eq:indi_inner_control_law}	
	\end{split}
\end{equation}
where $\bm{\dot{\Omega}}$ is a vector of commanded rotational rates for every rotor and $\tilde{T} = T - T_f$ being the thrust increment which is provided by the outer loop \acrshort{INDI}. The block diagram of the inner loop \acrshort{INDI} controller is presented in Figure \ref{fig:indi_inner_croped}.
\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=1\textwidth]{figs/indi_inner_croped.pdf}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Block diagram of the inner loop \acrshort{INDI} controller]{Block diagram of the inner loop \acrshort{INDI} controller. $A(z)$ is the actuator dynamics (see \ref{subsec:actuator_dynamics}) and $F(z)$ is the second-order filter (see \ref{subsec:filtering}).}
	\label{fig:indi_inner_croped}
\end{figure}
The inner \acrshort{INDI} controller which controls the angular acceleration is then augmented with two simple controllers which control the angular velocity $\bm{\omega}$ and the attitude $\bm{\eta}$ of the quadrotor (see Figure \ref{fig:pid_inner_croped}). Each of them consists of a single gain. Thus, the reference values of the \acrshort{INDI} part of the inner loop are provided by these two controllers. \begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=0.7\textwidth]{figs/pid_inner_croped.pdf}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Attitude controller as an augmentation of the inner loop \acrshort{INDI} controller]{Attitude controller as an augmentation of the inner loop \acrshort{INDI} controller.}
	\label{fig:pid_inner_croped}
\end{figure}

\subsection{Outer loop INDI for Quadrotor} \label{subsec: indi_outer}

@@

The linear acceleration $\bm{\dot{v}}_0$ which is obtained from accelerometer measurements is usually noisy. Applying the same filter as in the case of the inner loop derivation, the Equation \ref{eq:lin_momentum_indi_outer_lin} is given by 
\begin{equation}
	\begin{split}
		\bm{\dot{v}} = \bm{\dot{v}}_f +  m^{-1} \bm{G}(\Phi_0, \Theta_0, \Psi_0, T_0) (\bm{u} - \bm{u}_f)
		\label{eq:lin_momentum_indi_outer_lin_final}
	\end{split}
\end{equation}
Solving Equation \ref{eq:lin_momentum_indi_outer_lin_final} for $\bm{u}$ and substituting $\bm{\dot{v}}$ with the equivalent control variable $\bm{\nu}_{lin}$, the control law of the outer loop is obtained
\begin{equation}
	\begin{split}
		\bm{u_c} = \bm{u} = \bm{u}_f +  m \bm{G}^{-1}(\Phi_0, \Theta_0, \Psi_0, T_0) (\bm{\nu}_{lin} - \bm{\dot{v}}_f)
		\label{eq:indi_outer_control_law}
	\end{split}
\end{equation}
Figure \ref{fig:indi_outer_croped} shows the block diagram of the outer loop \acrshort{INDI}. 
\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=1\textwidth]{figs/indi_outer_croped.pdf}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Block diagram of the outer loop \acrshort{INDI} controller.]{Block diagram of the outer loop \acrshort{INDI} controller.}
	\label{fig:indi_outer_croped}
\end{figure}
Using the same approach as with the inner loop, the outer loop which controls linear acceleration is augmented with two closed loops consisting of two gains (see Figure \ref{fig:pid_outer_croped}). These gains represent two controllers which control linear velocity and position. 
\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=0.8\textwidth]{figs/pid_outer_croped.pdf}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Position controller as an augmentation of the outer loop \acrshort{INDI} controller]{Position controller as an augmentation of the outer loop \acrshort{INDI} controller.}
	\label{fig:pid_outer_croped}
\end{figure}

@@

The quadrotor used in this project was extended with a flow deck sensor. The flow deck itself is equipped with two sensors: the optical flow sensor which measures the visual motion and the \acrfull{ToF} sensor which measures the range to the next object along the body-fixed z-axis of the quadcopter. By employing these sensors, the flow deck can provide estimates of the current position and linear velocity of the aircraft.

\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=0.4\textwidth]{figs/crazyflie_drone.jpg}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Crazyflie 2.1 nano-quadrotor]{Crazyflie 2.1 nano-quadrotor \cite{bitcraze}.}
	\label{fig:crazyflie_drone}
\end{figure}

@@

In the framework of this thesis the Simulink model was also applied to better understand the controller structure and its characteristics. Based on the simulation it was determined which of the control effectiveness terms have a greater influence on the system response and which ones can be neglected. Additionally, the simulation model provided a good overview of potential computational problems and mistakes which could lead to drastically growing values and unstable behaviour of the system. To solve some of them, saturation blocks have been used  in different places.

@@

The core of the Crazyflie simulation is the model of the quadrotor dynamics. It is based on dynamic equations of motion presented in Chapter \ref{sec:eqs_motion}. Similarly to the section \ref{subsec: indi_outer}, it was assumed that the sum of the forces in the linear momentum equation is composed of three components, namely gravitational, propulsive and aerodynamic. The expressions of the gravity and propulsive forces have not been changed, so the Equations \ref{eq:Fg} and \ref{eq:Fp} were adopted for the simulation model. The thrust value $T$ was modelled as a linear function of rotational rates of the propellers 
\begin{equation}
	\begin{split}
		T = - k_F (\Omega_1 + \Omega_2 + \Omega_3 + \Omega_4)
		\label{eq:thrust}
	\end{split}
\end{equation}
where $k_F$ is the force constant of the rotors. In subsection \ref{subsec:parameter_estimation} it is shown how to estimate the thrust coefficient directly from the flight data. However, for the simulation model $k_F$ was approximated based on the assumption that the influence of the force constant of the rotors $k_F$ (considering the lever arm) is greater than the influence of the moment constant of the rotors $k_M$, and that the \textit{thrust to weight ratio} expression is $\frac{T}{W} \approx 2$.

@@

@@

%- To make the controller work a minimal knowledge of the system dynymics is need. Nevertherless G1, G2 still have to estimated accuratly because this 2 parameters do have inpact on the controller performance. \\
%- Plain step responses of the outer loop \\
%- Step responses with disturbance of the outer loop
